A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-10 survey.tex
Completed revision of Section III

The reader can find a  good presentation of task decomposition for AIs
playing RTS in~\cite{weber2011acs}. Although the previous task decomposition is common, a significant challenge is on designing architectures so that the individual AI techniques that address each of those tasks can communicate and effectively work together, resolving conflicts, prioritizing resources between them, etc. Section \ref{sec:integration} provides an overview of the task decompositions that state-of-the-art bots use.

@@

Reactive control, also called micro-management in RTS games, aims at maximizing the effectiveness of units, including simultaneous control of units of different types in complex battles on heterogeneous terrain.

% POTENTIAL FIELDS:
Potential fields (or influence maps) have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. A drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement Learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local optima.

@@

Additionally, there has been some interesting uses of reinforcement learning (RL) \cite{Sutton}: 
Wender and Watson \cite{WenderRL} evaluated the different major RL algorithms for (decentralized) micro-management, which perform all equally. Marthi et al. \cite{Marthi05} employ concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. Madeira et al. \cite{Madeira06} advocate the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. The actions spaces to explore is gigantic for real game setups. It requires to use the structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. Another approach is that proposed by Jaide and Mu{\~n}oz-Avila \cite{jaidee2012classq} through learning just one Q-function for each unit type, in order to cut down the search space.

Other approaches that aim at learning the parameters of an underlying model have also been explored. For example Ponsen and Spronck \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. The difficulty to work with multi-scale goals and plans is handled directly by case-based reasoning (CBR), which has been adapted for units behavior with continuous action models \cite{Molineaux08}, an integrated RL/CBR algorithm using continuous models, or with hybrid CBR/RL transfer learning \cite{CBR-RL}. Reactive planning \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration of low-level control. Synnaeve and Bessi\`{e}re \cite{SynnaeveMicroCig11} achieve hierarchical goals (coming from tactical decisions) integration through the addition of another sensory input corresponding to the goal's objective. Finally, evolutionary optimization by simulating fights can easily adapted to any parameter-dependent micro-management control model, as shown by \cite{OthmanSimu} which optimizes an AIIDE 2010 micro-management competition bot.

% Other:
Other research falling into reactive control has been performed in the field of cognitive science, where Wintermute et al. \cite{SORTS} have explored human-like attention models (with units grouping and vision of a unique screen location) for micro-management.

@@

@@

Thanks to the recent organization of international game AI competitions focused around the popular StarCraft game (see Section \ref{sec:competition}), several groups have been working on integrating many of the techniques described in the previous section into complete ``bots'', capable of playing complete StarCraft games. In this section we will overview some of the currently available top bots, and focus, first on the architecture being used to integrate all the techniques being used for the different aspects of the bot, and then on which subsets of techniques are used for each bot.

\subsection{RTS Bot Architectures}\label{sec:integration}

>>>>

The reader can find a  good presentation of task decomposition for AIs
playing RTS in~\cite{weber2011acs}. Although the previous task decomposition is common, a significant challenge is on designing architectures so that the individual AI techniques that address each of those tasks can communicate and effectively work together, resolving conflicts, prioritizing resources between them, etc. Section \ref{sec:bot} provides an overview of the task decompositions that state-of-the-art bots use.

@@

Reactive control %, also called micro-management in RTS games, 
aims at maximizing the effectiveness of units, including simultaneous control of units of different types in complex battles on heterogeneous terrain.

% POTENTIAL FIELDS:
Potential fields (or influence maps) have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

Additionally, there has been some interesting uses of reinforcement learning (RL) \cite{Sutton}: 
Wender and Watson \cite{WenderRL} evaluated the different major RL algorithms for (decentralized) micro-management, which perform all equally. Marthi et al. \cite{Marthi05} employ concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. Madeira et al. \cite{Madeira06} advocate the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. This is because the action space to explore is gigantic for real game setups. It requires exploiting the existing structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. Another approach is that proposed by Jaide and Mu{\~n}oz-Avila \cite{jaidee2012classq} through learning just one Q-function for each unit type, in order to cut down the search space. Other approaches that aim at learning the parameters of an underlying model have also been explored. For example Ponsen and Spronck \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. For example, evolutionary optimization by simulating fights can easily adapted to any parameter-dependent micro-management control model, as shown by \cite{OthmanSimu} which optimizes an AIIDE 2010 micro-management competition bot.

Finally, approaches based on game tree search are recently being explored for micro management. Churchill et al. \cite{churchill2012AIIDE} presented a variant of alpha-beta search capable of dealing with simultaneous moves and durative actions, which could handle reactive control for situations with up to eight versus eight units.

% Other:
Other research falling into reactive control has been performed in the field of cognitive science, where Wintermute et al. \cite{SORTS} have explored human-like attention models (with units grouping and vision of a unique screen location) for reactive control.

@@

A related problem is that of integrating reasoning at multiple levels of abstraction. Molineaux et al. \cite{Molineaux08} showed that the difficulty of working with multi-scale goals and plans can be handled directly by case-based reasoning (CBR), via an integrated RL/CBR algorithm using continuous models. Reactive planning \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration of low-level control. Synnaeve and Bessi\`{e}re \cite{SynnaeveMicroCig11} achieve hierarchical goals (coming from tactical decisions) integration through the addition of another sensory input corresponding to the goal's objective.

@@

Thanks to the recent organization of international game AI competitions focused around the popular StarCraft game (see Section \ref{sec:competition}), several groups have been working on integrating many of the techniques described in the previous section into complete ``bots'', capable of playing complete StarCraft games. In this section we will overview some of the currently available top bots.

%\subsection{RTS Bot Architectures}\label{sec:integration}
